---
title: "Imputation"
author: "Anna Luis"
date: "6 de diciembre de 2019"
output: html_document
---

Now all the data join is in this file data_train:
```{r}
library(readr)
data_train <- read_csv("data/data_train.csv")
```

```{r warning=FALSE, results="hide", message=FALSE} 
#Set Libraries:
library("readr")
library("dplyr")
library("lubridate")
```

Explicació sobre missing data. We have two options or do not use the variables containing it or imputate it.
We will start the imputation of the missing data with the weather data:

IMPUTATION

```{r}
library(readr)
weather_train <- read_csv("data/weather_train.csv")

```
```{r}
summary(is.na(weather_train))

```


There is quite a lot of missing data in this data set that can be relevant for the model.


To imputate the data two methods will performed:

First:
Taking into acount that the variables in this dataset are time depending. We will apply this method to all the varibles with missing data
1. When a missing value is found pick the two colindant values from the previous and later hour and make the mean.
2. If the two values are missing, pick the value of the same hour but from the previous day and next day and make as well the mean.
3. If the colindant days are also missing, pick the mean of the day, mean of the week and mean of the month.

```{r}
#let's create a table "imputation_table" with the new varibles for point 1 and 2 of this method.
imputation_table <- weather_train 

for(i in 3:9){
  column_name= colnames(weather_train)[i]
  col_previous= paste("previous",column_name, sep="_")
  col_next=paste("next",column_name, sep="_")
  col_last_day_same_hour=paste("last_day_same_hour",column_name, sep="_")
  col_next_day_same_hour=paste("next_day_same_hour",column_name, sep="_")

imputation_table<-imputation_table %>% group_by (site_id) %>% mutate(!!col_previous:=lag(!!as.name(column_name), n=1))
imputation_table<-imputation_table %>% group_by (site_id) %>% mutate(!!col_next:=lead(!!as.name(column_name), n=1))
imputation_table<-imputation_table %>% group_by (site_id) %>% mutate(!!col_last_day_same_hour:=lag(!!as.name(column_name), n=24))
imputation_table<-imputation_table %>% group_by (site_id) %>% mutate(!!col_next_day_same_hour:=lead(!!as.name(column_name), n=24))
 }
```

```{r}
imputation_table_mean_first <- imputation_table

for(i in 3:9 ){
 column_name= colnames(weather_train)[i]
 print(column_name)
 col_previous= paste("previous",column_name, sep="_")
 col_next=paste("next",column_name, sep="_")
 col_last_day_same_hour=paste("last_day_same_hour",column_name, sep="_")
 col_next_day_same_hour=paste("next_day_same_hour",column_name, sep="_")
 condition = is.na(imputation_table_mean_first)

 # first point of the model:
imputation_table_mean_first[condition[,column_name],column_name] = rowMeans(imputation_table_mean_first[condition[,column_name],c(col_previous, col_next)], na.rm= TRUE)

# recalibrate the condition for this variable:
condition = is.na(imputation_table_mean_first[,column_name])

# second point of the model:
imputation_table_mean_first[condition[,column_name],column_name] = rowMeans(imputation_table_mean_first[condition[,column_name],c( col_last_day_same_hour, col_next_day_same_hour)], na.rm= TRUE)

print(sum(is.na(imputation_table_mean_first[,column_name])))
}
```

After applying steps 1 and 2 there is still 4 variables (cloud_coverage, precip_depth_1_hr, sea_level_pressure and wind_direction) that has missing values. 
We will follow the logic and apply the 3rd step.


```{r}
#first create the variables:
imputation_table_mean_complete <- imputation_table_mean_first
imputation_table_mean_complete  <- imputation_table_mean_complete %>% mutate(timestamp_date = date(timestamp),
                          timestamp_month = month(timestamp),
                          timestamp_week = week(timestamp))


for(i in 3:9 ){
 column_name= colnames(weather_train)[i]
 print(column_name)
 col_mean_day= paste("mean_day",column_name, sep="_")
 col_mean_week=paste("mean_week",column_name, sep="_")
 col_mean_month=paste("mean_month",column_name, sep="_")
 col_mean_site=paste("mean_site",column_name, sep="_")
 col_mean_total=paste("mean_total",column_name, sep="_")
 
condition = is.na(imputation_table_mean_first)

day <- imputation_table_mean_complete %>%
  group_by(site_id,timestamp_date) %>%
  summarize(!!col_mean_day := mean(!!as.name(column_name), na.rm=TRUE))
 
imputation_table_mean_complete <- merge(imputation_table_mean_complete, day, by=c('timestamp_date', 'site_id'))

week <- imputation_table_mean_complete %>%
  group_by(site_id,timestamp_week) %>%
  summarize(!!col_mean_week := mean(!!as.name(column_name), na.rm=TRUE))

imputation_table_mean_complete <- merge(imputation_table_mean_complete, week, by=c('timestamp_week', 'site_id'))

month <- imputation_table_mean_complete %>%
  group_by(site_id,timestamp_month) %>%
  summarize(!!col_mean_month := mean(!!as.name(column_name), na.rm=TRUE))

imputation_table_mean_complete <- merge(imputation_table_mean_complete, month, by=c('timestamp_month', 'site_id'))

mean_site <- imputation_table_mean_complete %>%
  group_by(site_id) %>%
  summarize(!!col_mean_site := mean(!!as.name(column_name), na.rm=TRUE))

imputation_table_mean_complete <- merge(imputation_table_mean_complete, mean_site, by=c('site_id'))

#because there is still sites without any data of it, we will replaced by the mean of all the sites
imputation_table_mean_complete <- imputation_table_mean_complete %>%
 mutate(!!col_mean_total := mean(!!as.name(column_name), na.rm=TRUE))

}

```

```{r}
imputation_table_mean_final <- imputation_table_mean_complete

for(i in 3:9 ){
  column_name= colnames(weather_train)[i]
 print(column_name)
 
  for(j in 1:5){
 
mean_list <- c(paste("mean_day",column_name, sep="_"),
paste("mean_week",column_name, sep="_"),
paste("mean_month",column_name, sep="_"),
paste("mean_site",column_name, sep="_"),
paste("mean_total",column_name, sep="_"))
 
 
condition = is.na(imputation_table_mean_final)
print(mean_list[j])
print(sum(condition[,column_name]))

if(sum(condition[,column_name])>0){
  imputation_table_mean_final[condition[,column_name],column_name] = imputation_table_mean_final[condition[,column_name],mean_list[j]]}

print(sum(is.na(imputation_table_mean_final[,column_name])))
}}
```

```{r}
library(mice)
library(VIM)
```
The former implements a method to deal with missing data. The _mice_ package creates multiple imputations (replacement values) for multivariate missing data. The method is based on Fully Conditional Specification, where each incomplete variable is imputed by a separate model [1]. 

On the other hand, _VIM_ package stands for "Visualization and Imputation of Missing Values", and it introduces new tools for exploring the data and the structure of the missing or imputed values [2].
```{r}
imputation_table_mice <- imputation_table

imp <- mice(imputation_table_mice, seed = 7)

data.imp <- complete(imp, 1)

```

Now let's study which of the two imputation methods is better:
we will generate fake missing values and apply the two methodologies. Then compare using minimum square error as a metric, which method has a higher accuracy.
7000 rows are selected which represent 5% of the total.


```{r}
#duplicate the table
imputation_table_fake_na <- imputation_table
#number of values to replace:
N  <-  7000
#seed so the selection of the sample will always be the same:
set.seed(500)
#Aleatory sample without replacement:
ids <-sample(1:length(imputation_table_fake_na$air_temperature),N , replace = FALSE)
#Subsitute the rows selected by NA
imputation_table_fake_na[ids, "air_temperature"] <- NA 
#total na values:
sum(is.na(imputation_table_fake_na$air_temperature))

```

Now there are 7053 values, the 53 already existing plus the 7000 generated for testing the models.
Let's start by using the mean methodology:

```{r}
imputation_table_fake_na_mean <- imputation_table_fake_na

condition_is_fake_na = is.na(imputation_table_fake_na_mean$air_temperature)
imputation_table_fake_na_mean[condition_is_fake_na,"air_temperature"] = rowMeans(imputation_table_fake_na_mean[condition_is_fake_na,c("previous_temp","next_temp")], na.rm= TRUE)

condition_is_fake_na = is.na(imputation_table_fake_na_mean$air_temperature)
imputation_table_mean[condition_is_fake_na,"air_temperature"] = rowMeans(imputation_table_fake_na_mean[condition_is_fake_na,c("last_day_same_hour_temp","next_day_same_hour_temp")], na.rm= TRUE)
condition_is_fake_na = is.na(imputation_table_fake_na_mean$air_temperature)
sum(condition_is_fake_na)

```

and now the imputation using mice:



```{r}
imputation_table_fake_na_mice <- imputation_table_fake_na

imp <- mice(imputation_table_fake_na_mice, seed = 7)

imputation_table_fake_na_mice <- complete(imp, 1)

```

mean square error between the real values and the imputated.

```{r}

mean((imputation_table[ids,]$air_temperature-imputation_table_fake_na_mean[ids,]$air_temperature)^2, na.rm = TRUE)

```

```{r}

mean((imputation_table[ids,]$air_temperature-imputation_table_fake_na_mice[ids,]$air_temperature)^2, na.rm = TRUE)

```

0.38 // 19.18


The best method seemed to be the mean, so we will use this imputation.

For the rest of the variables we will apply in the weather data set that has nans

```{r}

mean((imputation_table[ids,]$air_temperature-imputation_table_fake_na_mice[ids,]$air_temperature)^2, na.rm = TRUE)

```
